---
title: "Agent Based Model"
author: "Candidate 13343"
date: "7/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load objects and libraries

```{r}
suppressMessages(library(stats))
suppressMessages(library(igraph))
suppressMessages(library(ggplot2))
```

```{r}
# read in the data frame of initial opinions from the LSS model 
initial_opinions <- readRDS("initial_opinions.RDS")

# get vector of opinions for each attack 
dens_vec_west <- subset(initial_opinions, attack == "Westminster")$fit_updated_scaled
dens_vec_manc <- subset(initial_opinions, attack == "Manchester Arena")$fit_updated_scaled
dens_vec_lond <- subset(initial_opinions, attack == "London Bridge")$fit_updated_scaled
dens_vec_fins <- subset(initial_opinions, attack == "Finsbury Park")$fit_updated_scaled
dens_vec_pars <- subset(initial_opinions, attack == "Parsons Green")$fit_updated_scaled

# estimate density function for each distribution
dens_west <- density(dens_vec_west)
dens_manc <- density(dens_vec_manc)
dens_lond <- density(dens_vec_lond)
dens_fins <- density(dens_vec_fins)
dens_pars <- density(dens_vec_pars)
```

************************************************************************************************************************

##### Set initial parameters

Function to create the network graph 

```{r}
# g <- watts.strogatz.game(1, 1000, 20, 0.35, loops = FALSE, multiple = FALSE)
# plot(g, vertex.label= NA, edge.arrow.size=0.02,vertex.size = 0.5, xlab = "Small world model")
```


```{r}
make_network <- function(num_agents, type = "small world") {
  
  if (type == "small world") {
    g <- sample_smallworld(1, num_agents, 10, 0.05)
  
  }
  
  if (type == "scale free") {
    
    g <- barabasi.game(num_agents, power = 1.2, m = NULL, out.dist = NULL, out.seq = NULL,
                       out.pref = FALSE, zero.appeal = 1, directed = FALSE,
                      algorithm ="psumtree", start.graph = NULL)
  }
  
  return(g)
}
```

```{r}
# network = setup_network(nx.watts_strogatz_graph, n, k=round(0.78*n), p=0.2)
```

Function to draw initial opinions for ABM based on results of LSS

```{r}
# get initial opinions- outside of function for reproducibility 
set.seed(123)

get_initial_opinions <- function(num_agents, attack_density, uniform_dist = FALSE) {

  if (uniform_dist == TRUE) {
    initial_opinions <- runif(num_agents, -1, 1) 
  } else {
    # from: https://stackoverflow.com/questions/17087312/generate-a-random-
    # number-from-a-density-object-or-more-broadly-from-a-set-of-nu
    initial_opinions <- sample(x = attack_density$x,
                               num_agents, 
                               prob = attack_density$y,
                               replace = TRUE) + rnorm(num_agents, 0, attack_density$bw)
  }
  
  return(initial_opinions)
}

initial_opinions_west <- get_initial_opinions(500, dens_west)
initial_opinions_manc <- get_initial_opinions(500, dens_manc)
initial_opinions_lond <- get_initial_opinions(500, dens_lond)
initial_opinions_fins <- get_initial_opinions(500, dens_fins)
initial_opinions_pars <- get_initial_opinions(500, dens_pars)
initial_opinions_unif <- runif(500, -1, 1)
```

Function to run the ABM model 

```{r}
# Function to run abm model
# Parameters: network graph - either small world or scale free
# Initial distribution of opinions - exploring 6 for each of the attacks + uniform distibution
# extremist threshold = -0.75 to align with LSS output 
# epsilon value - ranges from 0.1 to 0.9
run_abm <- function(network_graph, initial_opinions, num_iterations, extremist_threshold = -0.75, epsilon_val, verbose = FALSE) {

  # Number of agents in simulation
  num_agents <- vcount(network_graph)
  
  # Number of time steps to run 
  num_iterations <- num_iterations 
  
  # Threshold for which opinion is considered
  # "extremist", i.e., very anti-immigrant
  extremist_threshold <- extremist_threshold
  
  # Create data frame where each row is an agent, stores: 
  # - Current opinion [-1, 1]
  # - Epsilon (confidence bound for changing opinion)
  # - Extremist indicator (Boolean to indicate whether they are considered extremist)
  # - List of neighbors
  agent_df <- data.frame(
    agent_id = c(1:num_agents),
    initial_opinion = rep(NA, num_agents),
    current_opinion = rep(NA, num_agents), 
    epsilon = rep(NA, num_agents),
    extremist = rep(NA, num_agents),
    neighbors = rep(NA, num_agents)
    )
  
  # Create matrix where each row is an agent, each column a past opinion
  # (Easier to make separate from data frame)
  opinions_matrix <- matrix(
    NA,
    nrow = num_agents,
    ncol =  (num_iterations + 1))
  
  # Set initial value for past opinions, initial opinions, 
  # and current opinions
  agent_df$initial_opinion <- initial_opinions
  agent_df$current_opinion <- initial_opinions
  opinions_matrix[, 1] <- initial_opinions
    
  # Set initial extremist values based on current opinion
  agent_df$extremist <- ifelse(agent_df$current_opinion <= extremist_threshold, TRUE, FALSE)
  
  # Set epsilon values - 0.1 if extremist, otherwise epsilon
  agent_df$epsilon <- ifelse(agent_df$extremist == TRUE, 0.25, epsilon_val)
  
  # Initialise empty vector of neighbors 
  neighbors_list <- vector("list", num_agents)
  
  for (i in 1:nrow(agent_df)) {
    agent_neighbors <- as.list(neighbors(network_graph, i))
    neighbors_list[[i]] <- agent_neighbors
    }
  
  # Assign neighbors 
  agent_df$neighbors <- neighbors_list
  
  # Iterate over each time step
  for (i in 1:num_iterations){
  
    if (verbose == TRUE) {
      print(paste("Time step:", i))
    }
    
    # Select random indices of agents to update their opinion 
    # Starting with all of them each time
    agents_to_udpate <- sample(nrow(agent_df), floor(nrow(agent_df) / 3))
    
    # For each agent to update 
    for(j in 1:length(agents_to_udpate)) {
      
      # Get Agent A
      agent_a <- agents_to_udpate[j]
      opinion_agent_a <- agent_df$current_opinion[agent_a]

      # Select random neighbor to meet with - Agent B
      agent_b <- sample(agent_df$neighbors[agent_a][[1]], 1)[[1]]
      opinion_agent_b <- agent_df$current_opinion[agent_b]
      
      # If difference of opinions is less than epsilon
      if (abs(opinion_agent_a - opinion_agent_b) < agent_df$epsilon[agent_a]) {
        
        # Four scenarios: 
        # 1) A extremist, B extremist and more extreme than agent a
        # change opinion to B 
        if (agent_df$extremist[agent_a] == TRUE & (opinion_agent_b < opinion_agent_a)) {
          agent_df$current_opinion[agent_a] <- opinion_agent_b
        } 
        
        # 2) A extreme, B extreme but less than agent a
        # opinion remains the same
        if (agent_df$extremist[agent_a] == TRUE & (opinion_agent_b > opinion_agent_a)) {
          agent_df$current_opinion[agent_a] <- opinion_agent_a
        }
        
        # 3) A NOT extreme, B is extreme
        # pull closer to their opinion (Persuasive Arguments Theory and Social Comparison Theory)
        if (agent_df$extremist[agent_a] == FALSE & agent_df$extremist[agent_b] == TRUE) {
          agent_df$current_opinion[agent_a] <- 
            opinion_agent_a + (0.5 * (opinion_agent_b - opinion_agent_a)) + runif(1, -0.25, -0.05)
        }
        
        # 4) A is NOT extreme, B NOT extreme, opinions converge at rate mu = 0.5
        if (agent_df$extremist[agent_a] == FALSE & agent_df$extremist[agent_b] == FALSE) {
          agent_df$current_opinion[agent_a] <- opinion_agent_a + (0.5 * (opinion_agent_b - opinion_agent_a))
        }
        
      }
      
      # Update agent if it is now past extremist threshold  
      if (agent_df$current_opinion[agent_a] <= extremist_threshold) {
        agent_df$extremist[agent_a] <- TRUE
        agent_df$epsilon[agent_a] <- 0.1
      }
      
    }
    
    # Normalisation function 
    agent_df$current_opinion <- ifelse(agent_df$current_opinion > 1, 1, agent_df$current_opinion)
    agent_df$current_opinion <- ifelse(agent_df$current_opinion < -1, -1, agent_df$current_opinion)
    
    # update opinions matrix
    opinions_matrix[, (i + 1)] <-  agent_df$current_opinion
  }
  
  # return final opinions matrix and data frame 
  return(list(agent_df, opinions_matrix))
}
```

************************************************************************************************************************
# Run models for Each Attack 
- Each attack differs in the initial distribution of opinions 
- We are seeing how epsilon impacts the outcome (possibly the network as well, but starting with small world)

```{r}
num_agents <- 500
epsilon_vals <- seq(0.1, 0.9, 0.1)
g <- make_network(num_agents, type = "small world")
```

```{r cache=TRUE}
# separating because of the size of the output
# list to store 10 runs of each simulation
run_abm_iters <- function(initial_opinions) {
  
  abs_list_tot <- list()
  
  for (i in 1:5) {
  
    # list to store each simulation at the 9 epsilon values
    abm_list_eps <- list()
    
    for (j in 1:length(epsilon_vals)) {
      abm_output <- run_abm(network_graph = g,
                            initial_opinions = initial_opinions,
                            num_iterations = 2500,
                            epsilon_val = epsilon_vals[j])
  
      abm_list_eps[[j]] <- abm_output
      print(paste("epsilon round", i))
    }
  abs_list_tot[[i]] <- abm_list_eps
  print(paste("overall round", i))
  }
}
```

************************************************************************************************************************

Unpack the iteractinos and average for plotting 

```{r}
all_iter_agent_dfs <- data.frame()
all_iter_opinions_dfs <- data.frame()

# for each iteration
for (j in 1:5) {
  
  # get iteration - REPLACE WITH ATTACK
  iter_list <- abs_list_tot_pars[[j]]
  
  all_eps_agent_dfs <- data.frame()
  all_eps_opinions_df <- data.frame()
  
  # for each epsilon value
  for (i in 1:length(epsilon_vals)) {
  
    # get epsilons for that iteration
    eps_list <- iter_list[[i]]
    
    agent_df <- eps_list[[1]]
    opinions_matrix <- eps_list[[2]]
    
    # add epsilon value 
    agent_df$epsilon_value <- epsilon_vals[[i]]
  
    opinions_df <- as.data.frame(opinions_matrix) %>% 
      mutate(agent = c(1:nrow(opinions_matrix))) %>% 
      # Make agent first column
      relocate(agent, .before = V1) %>% 
      mutate(epsilon_value = epsilon_vals[[i]]) %>% 
      relocate(epsilon_value, .before = V1)
    
    # join to df of all epsilon values for each iteraion
    all_eps_agent_dfs <- rbind(all_eps_agent_dfs, agent_df)
    all_eps_opinions_df <- rbind(all_eps_opinions_df, opinions_df) 
  }
  
  # add iteraiton value 
  all_eps_agent_dfs$iteration <- j
  all_eps_opinions_df <- all_eps_opinions_df %>%
    mutate(iteration = j) %>% 
    relocate(iteration, .before = V1)
  
  # join to all iterations for this attack
  all_iter_agent_dfs <- rbind(all_iter_agent_dfs, all_eps_agent_dfs)
  all_iter_opinions_dfs <- rbind(all_iter_opinions_dfs, all_eps_opinions_df)
}
```

```{r}
# replace with attack name
opinions_df_pars <- all_iter_opinions_dfs %>% 
  group_by(agent, epsilon_value) %>% 
  summarise(across(everything(), list(mean)))

# remove iteration mean
opinions_df_pars <- opinions_df_pars[,-3]

# add attack 
opinions_df_pars <- opinions_df_pars %>% 
  mutate(attack = "Parsons Green") %>% 
  relocate(attack, .before = agent)

# rename timesteps (num iteration + 1)
colnames(opinions_df_pars)[4:ncol(opinions_df_pars)] <- c(1:2501)

# final agent_df for westminster - not used unless you want agent 
all_iter_agent_dfs_pars <- all_iter_agent_dfs

saveRDS(opinions_df_pars, "opinions_df_pars.RDS")
saveRDS(all_iter_agent_dfs_pars, "all_iter_agent_dfs_pars.RDS")
```

Hyperparameter grid 

```{r}
model_grid <- expand.grid(
  initial_distribution = c("Westminster", "Manchester Attack", "London Bridge",
                           "Finsbury Park", "Parsons Green", "Uniform"),
  epsilon = c(seq(0.1, 0.9, 0.1))
)
```

************************************************************************************************************************

#### Visualisations

##### Make agent trajectory graph - Include min and max epsilon for uniform distribution

```{r}
abm_list_eps <- list()

for (j in 1:length(epsilon_vals)) {
  abm_output <- run_abm(network_graph = g,
                        initial_opinions = initial_opinions_unif,
                        num_iterations = 2500,
                        epsilon_val = epsilon_vals[j])
  
  abm_list_eps[[j]] <- abm_output
}

unif_eps1_output <- abm_list_eps[[1]]
unif_eps9_output <- abm_list_eps[[9]]

agent_df_unif_eps1 <- unif_eps1_output[[1]]
opinions_matrix_unif_eps1 <- unif_eps1_output[[2]]

agent_df_unif_eps9 <- unif_eps9_output[[1]]
opinions_matrix_unif_eps9 <- unif_eps9_output[[2]]
```

```{r}
plot_trajectories <- function(agent_df, opinions_matrix, timestep_cutoff) {
  
  # iterations = number of timesteps 
  # num rounds = iterations + 1 (round 1 is timestep 0)
  num_agents <- nrow(opinions_matrix)
  num_rounds <- ncol(opinions_matrix)
  
  # Make opinions data frame for graphing
  opinion_df <- as.data.frame(opinions_matrix) %>% 
    mutate(agent = c(1:num_agents)) %>% 
    # Make agent first column
    relocate(agent, .before = V1)
  
  # Pivot df longer
  opinion_df_long <- opinion_df %>% 
    pivot_longer(cols = colnames(opinion_df)[2:length(colnames(opinion_df))],
                 names_to = "round", values_to = "opinion")
  
  opinion_df_long$round <- as.numeric(gsub("V", "", opinion_df_long$round))
  
  changed_agents <- (opinions_matrix[, 1] != opinions_matrix[, num_rounds])
  
  # Percentage of changed agents
  perc_changed <- sum(changed_agents)/num_agents
  
  # Percentage of extremists
  perc_extrmist <- sum(agent_df$extremist)/num_agents
  
  # discretize opinion space into 0.01 size bins and count 
  # the number of clusters (max 200 min 1)
  final_opinions <- round(opinion_df[, num_rounds], 2)
  num_clusters <- length(unique(final_opinions))
  
  # make plotting data frame for cutoff
  plot_df <- opinion_df_long %>% 
    filter(round <= timestep_cutoff)
  
  trajectory_plot <- plot_df %>% 
    ggplot(aes(x = round, 
               y = opinion, 
               group = agent)) + 
    # Create line of opinion over time
    geom_line(aes(colour = agent)) + 
    # adjust labels
    labs(title = "Agent opinion trajectories over time",
         subtitle = "ABM model of simiarity bias",
         caption = print(paste("Percent changed: ", round(perc_changed, 3),
                               "Percent extremist: ", round(perc_extrmist, 3),
                               "Number of clusters: ", num_clusters)),
         x = "Time Step",
         y = "Scaled opinion [-1, 1] ") +
    # adjust theme elements
    theme_bw() +
    theme(axis.title.x = element_text(margin = margin(10, 0, 0, 0)),
          axis.title.y = element_text(margin = margin(0, 10, 0, 0)),
          plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5),
          legend.position = "none",
          plot.margin = unit(c(0.25, 0.5, 0.25, 0.5), "cm"))
  
  return(trajectory_plot)
}
```

```{r}
plot_unif_eps1 <- plot_trajectories(agent_df_unif_eps1, opinions_matrix_unif_eps1, 750)
plot_unif_eps9 <- plot_trajectories(agent_df_unif_eps9, opinions_matrix_unif_eps9, 750)

ggsave(plot_unif_eps1, file = "plot_unif_eps1.png", width = 6, height = 8)
ggsave(plot_unif_eps9, file = "plot_unif_eps9.png", width = 6, height = 8)
```

# add uniform distribution to this 

```{r}
all_eps_agent_dfs <- data.frame()
all_eps_opinions_df <- data.frame()

# for each epsilon value
for (i in 1:length(epsilon_vals)) {

  # get epsilons for that iteration
  eps_list <- abm_list_eps[[i]]
  
  agent_df <- eps_list[[1]]
  opinions_matrix <- eps_list[[2]]
  
  # add epsilon value 
  agent_df$epsilon_value <- epsilon_vals[[i]]

  opinions_df <- as.data.frame(opinions_matrix) %>% 
    mutate(agent = c(1:nrow(opinions_matrix))) %>% 
    # Make agent first column
    relocate(agent, .before = V1) %>% 
    mutate(epsilon_value = epsilon_vals[[i]]) %>% 
    relocate(epsilon_value, .before = V1)
  
  # join to df of all epsilon values for each iteraion
  all_eps_agent_dfs <- rbind(all_eps_agent_dfs, agent_df)
  all_eps_opinions_df <- rbind(all_eps_opinions_df, opinions_df) 
}

# add attack 
opinions_df_unif <- all_eps_opinions_df %>% 
  mutate(attack = "Uniform initial distribution") %>% 
  relocate(attack, .before = agent)

# rename timesteps (num iteration + 1)
colnames(opinions_df_unif)[4:ncol(opinions_df_unif)] <- c(1:2501)
```

************************************************************************************************************************

Plot opinion shifts based on epsilon values, for each of the 5 attacks. 

************************************************************************************************************************

```{r}
# bind all attacks together 
opinions_df_all <- rbind(opinions_df_west, 
                         opinions_df_manc,
                         opinions_df_lond,
                         opinions_df_fins,
                         opinions_df_pars, 
                         opinions_df_unif)

# select first and final round values 
opinion_shift_plot_df <- opinions_df_all %>% 
  select(attack, agent, epsilon_value, "1", "2501") 

colnames(opinion_shift_plot_df)[4:5] <- c("initial_opinion", "final_opinion")

opinion_shift_plot_df <- opinion_shift_plot_df %>% 
  mutate(opinion_shift = final_opinion - initial_opinion)

# factorize variables for ordering 
opinion_shift_plot_df$epsilon_value <- as.factor(opinion_shift_plot_df$epsilon_value)

opinion_shift_plot_df <- transform(opinion_shift_plot_df,
                            attack = factor(attack, 
                                            levels = c("Westminster", "Manchester Arena", "London Bridge",
                                                       "Finsbury Park", "Parsons Green", "Uniform initial distribution")))

opinion_shift_plot <- opinion_shift_plot_df %>% 
    ggplot() + 
    geom_boxplot(aes(x = epsilon_value, y = opinion_shift)) + 
    # adjust labels
    labs(title = "Opinion shift for each epsilon value",
         x = "Epsilon value",
         y = "Opinion shift (final - initial opinion)") +
    # adjust theme elements
    theme_bw() +
    theme(axis.title.x = element_text(size = 12, margin = margin(10, 0, 0, 0), face = "bold"),
          axis.title.y = element_text(size = 12, margin = margin(0, 10, 0, 0), face = "bold"),
          plot.title = element_text(size = 18, hjust = 0.5),
          axis.text = element_text(size = 10, hjust = 0.5),
          plot.margin = unit(c(0.25, 0.5, 0.25, 0.5), "cm")) +
  facet_wrap(~ attack, scales = "free_y", nrow = 3, ncol = 2) +
        theme(strip.text.x = element_text(size = 12))


ggsave(opinion_shift_plot, file = "opinion_shift_plot.png", height = 14, width = 12)
```


************************************************************************************************************************

Create table looking at changes from initial to final distribution 

```{r}
abm_table <- opinions_df_all %>% 
  select(attack, agent, epsilon_value, "1", "2501") 

colnames(abm_table)[4:5] <- c("initial_opinion", "final_opinion")

abm_table <- abm_table %>% 
  mutate(opinion_shift = final_opinion - initial_opinion)

# get extremists
extreme_table <- abm_table %>% 
  mutate(init_extremist = ifelse(initial_opinion <= -0.75, TRUE, FALSE)) %>% 
  mutate(final_extremist = ifelse(final_opinion <= -0.75, TRUE, FALSE))

extreme_table <- ungroup(extreme_table)

init_extreme <- extreme_table %>% 
  group_by(attack, epsilon_value) %>% 
  summarise(perc_init_extreme = sum(init_extremist) / n()) %>% 
  ungroup()

final_extreme <- extreme_table %>% 
  group_by(attack, epsilon_value) %>% 
  summarise(perc_final_extreme = sum(final_extremist) / n()) %>% 
  ungroup()

# get number of clusters
  summarise(across(everything(), list(mean)))

abm_table <- ungroup(abm_table)

abm_table_final <- abm_table %>% 
  group_by(attack, epsilon_value) %>% 
  summarise(across(everything(), list(mean)))  

abm_table_final <- abm_table_final[,-3]

abm_table_final[, -c(1:2)] <- round(abm_table_final[, -c(1:2)], 4)

# add in extrmist 
abm_table_final <- abm_table_final %>% 
  left_join(init_extreme, by = c("attack", "epsilon_value")) %>% 
  left_join(final_extreme, by = c("attack", "epsilon_value")) 

abm_table_paper <- abm_table_final %>% 
  filter(epsilon_value %in% c(0.1, 0.5, 0.9)) 

write.csv(abm_table_paper,'abm_table_paper.csv')
```

************************************************************************************************************************

##### Final opinion as function of initial opinion

```{r}
plot_initial_final_opinions <- function(agent_df, opinions_matrix) {

  # iterations = number of timesteps 
  # num rounds = iterations + 1 (round 1 is timestep 0)
  num_agents <- nrow(opinions_matrix)
  num_rounds <- ncol(opinions_matrix)
  
  # Make opinions data frame for graphing
  opinions_df <- data.frame(initial_opinion = opinions_matrix[, 1],
                            final_opinion = opinions_matrix[, num_rounds],
                            extremist = agent_df$extremist)
  
  opinions_df <- opinions_df %>% 
    mutate(became_extremist = ifelse(initial_opinion > -0.9 & 
                                       final_opinion <= -0.9, TRUE, FALSE))
  
  
  opinion_function_plot <- opinions_df %>% 
    ggplot(aes(x = initial_opinion,
               y = final_opinion)) +
    geom_point(aes(colour = became_extremist)) + 
    # adjust labels
    labs(title = "Final opinion as a function of initial opinion",
         x = "Initial Opinion",
         y = "Final Opinion") +
    # adjust theme elements
    theme_minimal() +
    theme(axis.title.x = element_text(margin = margin(10, 0, 0, 0)),
          axis.title.y = element_text(margin = margin(0, 10, 0, 0)),
          legend.position = "none",
          plot.margin = unit(c(0.25, 0.5, 0.25, 0.5), "cm"))
  
  return(opinion_function_plot)
} 
```

##### Mean and median opinions over time

```{r}
plot_mean_median <- function(agent_df, opinions_matrix) {
  
  # iterations = number of timesteps 
  # num rounds = iterations + 1 (round 1 is timestep 0)
  num_agents <- nrow(opinions_matrix)
  num_rounds <- ncol(opinions_matrix)
  
  opinion_df <- data.frame(t(opinions_matrix))
  
  timestep_mean <- rowMeans(opinion_df)
  timestep_median <- rowMedians(as.matrix(opinion_df))
  
  opinion_df$timestep_mean <- timestep_mean
  opinion_df$timestep_median <- timestep_median
  
  opinion_df <- opinion_df %>% 
    mutate(timestep = c(1:nrow(opinion_df)))
  
  mean_median_plot <- opinion_df %>% 
    ggplot(aes(x = timestep,
               y = timestep_mean)) +
    geom_line(colour = "blue") +
    geom_line(aes(x = timestep,
                  y = timestep_median)) +
    geom_hline(yintercept = 0,
               alpha = 0.4) +
    # adjust labels
    labs(title = "Median and mean opinion over time",
         x = "Time step",
         y = "Opinion Scale") +
    # adjust theme elements
    theme_minimal() +
    theme(axis.title.x = element_text(margin = margin(10, 0, 0, 0)),
          axis.title.y = element_text(margin = margin(0, 10, 0, 0)),
          legend.position = "none",
          plot.margin = unit(c(0.25, 0.5, 0.25, 0.5), "cm"))
    
  return(mean_median_plot)
}
```



```{r}
plot_dist_ind <- function(agent_df, opinions_matrix){

  num_agents <- nrow(opinions_matrix)
  num_rounds <- ncol(opinions_matrix)
  
  # Make opinions data frame for graphing
  opinion_df <- as.data.frame(opinions_matrix) %>% 
    mutate(agent = c(1:num_agents)) %>% 
    # Make agent first column
    relocate(agent, .before = V1)
  
  # Pivot df longer
  opinion_df_long <- opinion_df %>% 
    pivot_longer(cols = colnames(opinion_df)[2:length(colnames(opinion_df))],
                 names_to = "round", values_to = "opinion")
    
  opinion_df_long$round <- as.numeric(gsub("V", "", opinion_df_long$round))
  
  plot_df <- opinion_df_long %>% 
    filter(round == 1 | round == num_rounds) %>% 
    mutate(distribution = ifelse(round == 1, "initial", "final"))
  
  plot_dens_means <- plot_df %>% 
    group_by(distribution) %>% 
    summarise(mean_opinion = mean(opinion))
  
  
  dist_plot <- plot_df %>% 
    ggplot(aes(x = opinion, 
               fill = distribution)) + 
    geom_density(alpha = 0.3) +
    geom_vline(data = plot_dens_means, 
               aes(xintercept = mean_opinion,
                   colour = distribution),
               linetype = "dashed") + 
    # adjust labels
    labs(title = "Density distribution of initial and final opinions",
         x = "Scaled opinion value",
         y = "Frequency") +
    # adjust theme elements
    theme_minimal() +
    theme(axis.title.x = element_text(margin = margin(10, 0, 0, 0)),
          axis.title.y = element_text(margin = margin(0, 10, 0, 0)),
          plot.title = element_text(hjust = 0.5),
          plot.margin = unit(c(0.25, 0.5, 0.25, 0.5), "cm"))
    # facet_wrap(~ attack, scale = "free_y", nrow = 2, ncol = 3)
    
  return(dist_plot)
}

```
```{r}
plot_dist_all_eps <- function(abm_list) {
  
  plot_df_all_eps <- data.frame()
  plot_dens_means_all_eps <- data.frame()
  
  for (i in 1:length(abm_list)) {
    
    abm <- abm_list[[i]]
    agent_df <- abm[[1]]
    opinions_matrix <- abm[[2]]
    
    # iterations = number of timesteps 
    # num rounds = iterations + 1 (round 1 is timestep 0)
    num_agents <- nrow(opinions_matrix)
    num_rounds <- ncol(opinions_matrix)
    
    # Make opinions data frame for graphing
    opinion_df <- as.data.frame(opinions_matrix) %>% 
      mutate(agent = c(1:num_agents)) %>% 
      # Make agent first column
      relocate(agent, .before = V1)
    
    # Pivot df longer
    opinion_df_long <- opinion_df %>% 
      pivot_longer(cols = colnames(opinion_df)[2:length(colnames(opinion_df))],
                   names_to = "round", values_to = "opinion")
      
    opinion_df_long$round <- as.numeric(gsub("V", "", opinion_df_long$round))
    
    plot_df <- opinion_df_long %>% 
      filter(round == 1 | round == num_rounds) %>% 
      mutate(distribution = ifelse(round == 1, "initial", "final")) %>% 
      mutate(epsilon_val = epsilon_vals[[i]])

    plot_dens_means <- plot_df %>% 
      group_by(distribution) %>% 
      summarise(mean_opinion = mean(opinion)) %>% 
      mutate(epsilon_val = epsilon_vals[[i]])

    
    plot_df_all_eps <- rbind(plot_df_all_eps, plot_df)
    plot_dens_means_all_eps <- rbind(plot_dens_means_all_eps, plot_dens_means)
  }
  
  plot_df_all_eps$epsilon_val <- as.factor(plot_df_all_eps$epsilon_val)
  plot_dens_means_all_eps$epsilon_val <- as.factor(plot_dens_means_all_eps$epsilon_val)
  
  
  dist_plot <- plot_df_all_eps %>% 
    ggplot(aes(x = opinion, 
               fill = distribution)) + 
    geom_density(alpha = 0.3) +
    geom_vline(data = plot_dens_means_all_eps, 
               aes(xintercept = mean_opinion,
                   colour = distribution),
               linetype = "dashed") + 
    # adjust labels
    labs(title = "Density distribution of initial and final opinions",
         x = "Scaled opinion value",
         y = "Frequency") +
    # adjust theme elements
    theme_minimal() +
    theme(axis.title.x = element_text(margin = margin(10, 0, 0, 0)),
          axis.title.y = element_text(margin = margin(0, 10, 0, 0)),
          plot.title = element_text(hjust = 0.5),
          plot.margin = unit(c(0.25, 0.5, 0.25, 0.5), "cm")) +
  facet_wrap(~ epsilon_val, scale = "free_y", nrow = 3, ncol = 3)
    
  return(dist_plot)
}
```

```{r}
plot_factors <- function(agent_df, opinions_matrix) {

  num_agents <- nrow(opinions_matrix)
  num_rounds <- ncol(opinions_matrix)
  
  # Make opinions data frame for graphing
  opinion_df <- as.data.frame(opinions_matrix) %>% 
    mutate(agent = c(1:num_agents)) %>% 
    # Make agent first column
    relocate(agent, .before = V1)
  
  # Pivot df longer
  opinion_df_long <- opinion_df %>% 
    pivot_longer(cols = colnames(opinion_df)[2:length(colnames(opinion_df))],
                 names_to = "round", values_to = "opinion")
    
  opinion_df_long$round <- as.numeric(gsub("V", "", opinion_df_long$round))
  
  opinion_df_long <- opinion_df_long %>% 
    mutate(extremity_factor = ifelse(opinion < 0,
                                     ifelse(opinion < -0.25, 
                                            ifelse(opinion < -0.75, "strongly_negative", "negative"),
                                            "slightly_negative"),
                                     ifelse(opinion > 0.25,
                                            ifelse(opinion > 0.75, "strongly_positive", "positive"),
                                            "slightly_positive")))
  
  plot_df <- opinion_df_long %>% 
    filter(round == 1 | round == num_rounds) 
  
  plot_df$round <- as.factor(plot_df$round)
  plot_df$extremity_factor <- factor(plot_df$extremity_factor,
                                     levels = c("strongly_positive", "positive", "slightly_positive",
                                                "slightly_negative", "negative", "strongly_negative"))
  plot_df_num <- plot_df %>% 
    group_by(round, extremity_factor) %>% 
    summarise(n = n()) %>% 
    mutate(freq = n / sum(n))
  
  factor_plot <- ggplot(arrange(plot_df_num, extremity_factor),
                        aes(x = round, 
                            y = freq,
                            fill = extremity_factor)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = c("strongly_positive" = "dodgerblue4",
                                 "positive" = "dodgerblue1",
                                 "slightly_positive" = "skyblue1",
                                 "slightly_negative" = "lightcoral",
                                 "negative" = "indianred",
                                 "strongly_negative" = "firebrick4"))
  
  return(factor_plot)
}
```
